---
title: "Estimating Tropical Principal Components using Metropolis Hasting Algorithm"
author: "Qiwen Kang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Introduction

This note is wrote for showing how to use R code implementing MCMC method in tropical space. All the code included in this note could be found on https://github.com/QiwenKang/tropMCMC.git.

## Getting started

First of all, we need to setup an proper environment and load all the packages we need.

```{r eval=FALSE}
library(ape)
library(phangorn)
library(tropMCMC)
```

## MCMC method

After setting up enviroment, we need to read in an alignment. In our paper, we used lungfish data set, which you can download from http://www.mas.ncl.ac.uk/~ntmwn/geophytterplus/, as an example. The function `read.tree` is used to read in an tree.

```{r eval=FALSE}
trees_ori <- read.tree('./data/lungfish.txt')
```

Once we have read in all the raw trees, we need to declare some variable name meanwhile. `n` is the number of leaves; `pcs` gives how many principle components would be considered; `to` is tip labels of the raw trees and ordered by names; `N` is the number of trees in whole data set.

```{r eval=FALSE}
n <- length(trees_ori[[1]]$tip.label)
N <- length(trees_ori)
pcs <- 3
to <- trees_ori[[1]]$tip.label
```
Until now, we have already known some basic information about our data set. To begin our method, we should extract distance matrix from each tree and transfer it to a vector format. This could be done using `disMat` function. We need to state outgroups here as well. Variable `D_all` is just a matrix format of `distVec_all` with number of rows equals to `pcs` and number of columns equals to `N`.  

```{r eval=FALSE}
distVec_all <- distMat(trees_ori, tipOrder = to, outgp = c("Leucoraja","Scyliorhin","Callorhinc"))
D_all <- matrix(unlist(distVec_all), ncol = N) 
```
Now, we can begin our MCMC method. `sumsValue`, `comb_list`, `points_list` and `trop_list` are four variables to save the output of our method. Specifically, `sumsValue` is the sum of tropical distances; `points_list` includes all the projected points on tropical space of each tree; `comb_list` includes the index of trees which three trees have the minimum sum of tropical distances; `trop_list` contains the tropical distance between two points.

```{r eval=FALSE}
tropMCMC(distVect_all = distVec_all, N = N, nr= 5, pcs = pcs)
```
